---
title: "Heuristic Analysis"
author: "Arun K Viswanathan"
date: "Feb 18th 2017"
output:
  pdf_document:
    highlight: zenburn
    toc: no
  html_document:
    theme: united
    toc: yes
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)
```

```{r load-libraries, echo=FALSE, include=FALSE}
library(ggplot2)
```

```{r data, echo=FALSE, include=FALSE}
df <- data.frame(Heuristic = character(),
                 Player = character(), 
                 Opponent = character(), 
                 Wins = numeric(),
                 Losses = numeric(),
                 stringsAsFactors = TRUE)

df <- rbind(df, data.frame(Heuristic="Ratio",Player="ID_Improved",Opponent="Random",Wins=20,Losses=0))
df <- rbind(df, data.frame(Heuristic="Ratio",Player="ID_Improved",Opponent="MM_Null",Wins=17,Losses=3))
df <- rbind(df, data.frame(Heuristic="Ratio",Player="ID_Improved",Opponent="MM_Open",Wins=10,Losses=10))
df <- rbind(df, data.frame(Heuristic="Ratio",Player="ID_Improved",Opponent="MM_Improved",Wins=10,Losses=10))
df <- rbind(df, data.frame(Heuristic="Ratio",Player="ID_Improved",Opponent="AB_Null",Wins=12,Losses=8))
df <- rbind(df, data.frame(Heuristic="Ratio",Player="ID_Improved",Opponent="AB_Open",Wins=8,Losses=12))
df <- rbind(df, data.frame(Heuristic="Ratio",Player="ID_Improved",Opponent="AB_Improved",Wins=7,Losses=13))

df <- rbind(df, data.frame(Heuristic="Ratio",Player="Student",Opponent="Random",Wins=18,Losses=2))
df <- rbind(df, data.frame(Heuristic="Ratio",Player="Student",Opponent="MM_Null",Wins=15,Losses=5))
df <- rbind(df, data.frame(Heuristic="Ratio",Player="Student",Opponent="MM_Open",Wins=11,Losses=9))
df <- rbind(df, data.frame(Heuristic="Ratio",Player="Student",Opponent="MM_Improved",Wins=11,Losses=9))
df <- rbind(df, data.frame(Heuristic="Ratio",Player="Student",Opponent="AB_Null",Wins=13,Losses=7))
df <- rbind(df, data.frame(Heuristic="Ratio",Player="Student",Opponent="AB_Open",Wins=7,Losses=13))
df <- rbind(df, data.frame(Heuristic="Ratio",Player="Student",Opponent="AB_Improved",Wins=12,Losses=8))

df <- rbind(df, data.frame(Heuristic="Avg_Reduction",Player="ID_Improved",Opponent="Random",Wins=20,Losses=0))
df <- rbind(df, data.frame(Heuristic="Avg_Reduction",Player="ID_Improved",Opponent="MM_Null",Wins=14,Losses=6))
df <- rbind(df, data.frame(Heuristic="Avg_Reduction",Player="ID_Improved",Opponent="MM_Open",Wins=11,Losses=9))
df <- rbind(df, data.frame(Heuristic="Avg_Reduction",Player="ID_Improved",Opponent="MM_Improved",Wins=10,Losses=10))
df <- rbind(df, data.frame(Heuristic="Avg_Reduction",Player="ID_Improved",Opponent="AB_Null",Wins=12,Losses=8))
df <- rbind(df, data.frame(Heuristic="Avg_Reduction",Player="ID_Improved",Opponent="AB_Open",Wins=8,Losses=12))
df <- rbind(df, data.frame(Heuristic="Avg_Reduction",Player="ID_Improved",Opponent="AB_Improved",Wins=8,Losses=12))

df <- rbind(df, data.frame(Heuristic="Avg_Reduction",Player="Student",Opponent="Random",Wins=12,Losses=8))
df <- rbind(df, data.frame(Heuristic="Avg_Reduction",Player="Student",Opponent="MM_Null",Wins=6,Losses=14))
df <- rbind(df, data.frame(Heuristic="Avg_Reduction",Player="Student",Opponent="MM_Open",Wins=4,Losses=16))
df <- rbind(df, data.frame(Heuristic="Avg_Reduction",Player="Student",Opponent="MM_Improved",Wins=4,Losses=16))
df <- rbind(df, data.frame(Heuristic="Avg_Reduction",Player="Student",Opponent="AB_Null",Wins=7,Losses=13))
df <- rbind(df, data.frame(Heuristic="Avg_Reduction",Player="Student",Opponent="AB_Open",Wins=2,Losses=18))
df <- rbind(df, data.frame(Heuristic="Avg_Reduction",Player="Student",Opponent="AB_Improved",Wins=3,Losses=17))

df <- rbind(df, data.frame(Heuristic="Reachable",Player="ID_Improved",Opponent="Random",Wins=18,Losses=2))
df <- rbind(df, data.frame(Heuristic="Reachable",Player="ID_Improved",Opponent="MM_Null",Wins=19,Losses=1))
df <- rbind(df, data.frame(Heuristic="Reachable",Player="ID_Improved",Opponent="MM_Open",Wins=7,Losses=13))
df <- rbind(df, data.frame(Heuristic="Reachable",Player="ID_Improved",Opponent="MM_Improved",Wins=10,Losses=10))
df <- rbind(df, data.frame(Heuristic="Reachable",Player="ID_Improved",Opponent="AB_Null",Wins=10,Losses=10))
df <- rbind(df, data.frame(Heuristic="Reachable",Player="ID_Improved",Opponent="AB_Open",Wins=6,Losses=14))
df <- rbind(df, data.frame(Heuristic="Reachable",Player="ID_Improved",Opponent="AB_Improved",Wins=12,Losses=8))

df <- rbind(df, data.frame(Heuristic="Reachable",Player="Student",Opponent="Random",Wins=17,Losses=3))
df <- rbind(df, data.frame(Heuristic="Reachable",Player="Student",Opponent="MM_Null",Wins=16,Losses=4))
df <- rbind(df, data.frame(Heuristic="Reachable",Player="Student",Opponent="MM_Open",Wins=10,Losses=10))
df <- rbind(df, data.frame(Heuristic="Reachable",Player="Student",Opponent="MM_Improved",Wins=6,Losses=14))
df <- rbind(df, data.frame(Heuristic="Reachable",Player="Student",Opponent="AB_Null",Wins=12,Losses=8))
df <- rbind(df, data.frame(Heuristic="Reachable",Player="Student",Opponent="AB_Open",Wins=11,Losses=9))
df <- rbind(df, data.frame(Heuristic="Reachable",Player="Student",Opponent="AB_Improved",Wins=4,Losses=16))

```

The following custom heuristics are implemented in game play for isolation:

1. Ratio 
- In this heuristic, the score is computed as the ratio of the number of moves for the agent to the number of moves for the
opponent player in teh given state. Higher ratios imply better positions for the agent.

2. Average Reduction
- In this heuristic, the average reduction in the number of moves for the opponent, by moves of the agent, is used as a
measure of the score.

3. Reachability
- This heuristic uses the fraction of the blank spaces in the board that can be reached by the agent's next moves as a measure of the score.

The plot below compares the win percentages for ID_Improved and the Student's `CustomPlayer` against the different opponents, with each of the three heuristics.

```{r plot, echo=FALSE, include=TRUE}
df["WinPercent"] <- df["Wins"]/(df["Wins"]+df["Losses"])
plot <- ggplot(df, aes(x = Opponent, y = WinPercent, group = Player, col = Player)) + 
  geom_point(size = 2, shape = 21) + 
  geom_line(linetype = "solid", size = 1) + 
  facet_grid(. ~ Heuristic) +
  ylim(0, 1) +
  ylab("Win Percentage") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot
```

As seen in the plot, the ratio heuristic appears to be the best of three heuristics.